A Switching Planner for Combined Task and Observation Planning

### 2011-02-04 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
shrinking and growing, all at the same time.  Not sure how we are going to get this down to 6 pages yet.

From an automated planning perspective, the problem of practical
mobile robot control poses important and contrary challenges.
%%
On the one hand, planning and execution monitoring must be
lightweight, robust, timely, and should span the lifetime of the
robot. Those processes must seamlessly accommodate exogenous events,
changing objectives, and the underlying unpredictability of the
environment.
%%
On the other hand, robot planning should perform computationally
expensive reasoning about contingencies, and possible revisions of
subjective belief according to quantitatively modelled uncertainty in
acting and sensing.

In this paper we address these challenges, developing a continual
planner that switches between fast sequential and expensive
decision-theoretic planning. Given a POMDP model of the environment,
sequential planning is used to compute an initial deterministic
sequential plan and complementary runtime evolution of the decision
process. That plan is executed until testing of an assumption about
the runtime state is scheduled. At that point a contingent session
performs some sensing which determines whether or not the sequential
session should be allowed to continue, or otherwise acts to achieve
the overall objectives.

>>>>

%% From an automated planning perspective, the problem of practical
%% mobile robot control poses important and contrary challenges.
%% %%
%% On the one hand, planning and execution monitoring must be
%% lightweight, robust, timely, and should span the lifetime of the
%% robot. Those processes must seamlessly accommodate exogenous events,
%% changing objectives, and the underlying unpredictability of the
%% environment.
%% %%
%% On the other hand, robot planning should perform computationally
%% expensive reasoning about contingencies, and possible revisions of
%% subjective belief according to quantitatively modelled uncertainty in
%% acting and sensing.

For use in an integrated continual planning and execution monitoring
system on a mobile robot platform, we developed a base planner that
switches between fast sequential and expensive decision-theoretic
planning. Given a POMDP model of the environment, sequential planning
is used to compute an initial deterministic sequential plan and
complementary runtime evolution of the decision process. That plan is
executed until testing of an assumption about the runtime state is
scheduled. At that point a contingent session performs some sensing
which determines whether or not the sequential session should be
allowed to continue, or otherwise acts to achieve the overall
objectives.

### 2011-02-04 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
shrinking and growing, all at the same time.  Not sure how we are going to get this down to 6 pages yet.

There have been a number of recent papers on
planning under uncertainty using systems that were intended for
sequential planning in deterministic problems,  for example,
\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
probabilistic track of the 2004 International Planning Competition.
In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
plans and corresponding execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures. The outcome is typically a more
robust sequential plan~\cite{yoon:etal:2008}, or contingent
plan~\cite{majercik:2006}.

>>>>

%%  and
%% using various techniques to manage the large state
%% space

%% \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
%% approach in a

Addressing task and observation planning specifically, there have been
a number of recent developments where the underlying problem is
modelled as a POMDP.
%%
For vision algorithm
selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit a
natural hierarchical decomposition of the underlying POMDP into
subproblems where decision-theoretic planning is reasonably
fast. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
represent a preference elicitation problem as a POMDP and take
advantage of symmetry in the belief space ---essentially the idea that
it does not matter what the value of the variable you are trying to
observe is--- to exponentially shrink the state space. Although we
have been actively exploring the \citeauthor{doshi08:pref_elic}
approach, those exploitable structures are not present in problems we
have considered so far due to the task planning requirement. 
%%
%%
Finally,
our approach is in a similar vein to the more classical {\em
dual-mode} control~\cite{cassandra96actingunder}, however in our case
entropy reduction is targeted by planning in an abstract process which
is informed by one execution trace ---computed by a {\em classical}
planner--- and the underlying belief-state.

More generally, there has been much recent work on scaling POMDP
solution procedures to medium-sized
instances~\cite{shani:etal:08,KurHsu08}. In that
setting \citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010})
recently addressed an inefficiency of offline point-based techniques
in problems with medium length planning horizons, however their
approach take tens of minutes to good plans, and only scales to tens
of thousands of states. In the case of general domain-independent
factored systems, the state-of-the-art scales to relatively small
problems with $2^{22}$ states~\cite{shani:etal:2008}. At their limit,
these procedures take over an hour to converge, and $\sim10$ seconds
on average to perform a single Bellman backup.  For classes of POMDP
that feature exploitable structures (e.g., no actions with negative
effects), problems with as many as $10^{30}$ states can be targeted by
offline procedures~\cite{brunskill:russell:2010}. Moving someway
towards addressing real-time decision making, recent online POMDP
solution procedures have been developed which can exploit highly
approximate value functions -- typically computed using a point-based
procedure -- and heuristics in forward
search~\cite{ross:etal:2008}. Their applicability in our setting is
limited, firstly because they only scale to smaller problems, with
thousands of states, and also due to the large amount of
\emph{problem-specific} offline processing that might be required to get useful
search guidance. A {\em very} recent and promising online approach for
large POMDPs employs Monte-Carlo sampling to break the curse of
dimensionality in situations where goal reachability is {\em easily}
determined~\cite{silver:veness:2010}.

%% Although we suppose it an interesting
%% item for future work to pursue that direction, it should be noted that
%% ease of goal reachability is not guaranteed in the problems we face,
%% and is certainly not a property to be assumed in domain independent
%% planning.

In the direction of leveraging {\em classical} systems/approaches for
planning under uncertainty, the most highlighted system to date has
been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
the probabilistic track of the 2004 International Planning
Competition.  In the continual paradigm, \system{FFR$_a$}
uses \system{FF} to compute sequential plans and execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
%% robust sequential plan, or contingent
%% plan~\cite{majercik:2006}.
